Predicting the Quality of CDRH3 Antibody Loop Structural Models

### 2022-10-05 paper.tex
Changed paper to standard LaTeX and split lines

%% 
%% Copyright 2007-2020 Elsevier Ltd
%% 
%% This file is part of the 'Elsarticle Bundle'.
%% ---------------------------------------------
%% 
%% It may be distributed under the conditions of the LaTeX Project Public
%% License, either version 1.2 of this license or (at your option) any
%% later version.  The latest version of this license is in
%%    http://www.latex-project.org/lppl.txt
%% and version 1.2 or later is part of all distributions of LaTeX
%% version 1999/12/01 or later.
%% 
%% The list of all files belonging to the 'Elsarticle Bundle' is
%% given in the file `manifest.txt'.
%%

%% Template article for Elsevier's document class `elsarticle'
%% with numbered style bibliographic references
%% SP 2008/03/01
%%
%% 
%%
%% $Id: elsarticle-template-num.tex 190 2020-11-23 11:12:32Z rishi $
%%
%%
%Path relative to the .tex file containing the \includegraphics command

\documentclass[preprint,12pt]{elsarticle}
\usepackage{graphicx}
\graphicspath{ {./Images/} }
%% Use the option review to obtain double line spacing
%% \documentclass[authoryear,preprint,review,12pt]{elsarticle}

%% Use the options 1p,twocolumn; 3p; 3p,twocolumn; 5p; or 5p,twocolumn
%% for a journal layout:
%% \documentclass[final,1p,times]{elsarticle}
%% \documentclass[final,1p,times,twocolumn]{elsarticle}
%% \documentclass[final,3p,times]{elsarticle}
%% \documentclass[final,3p,times,twocolumn]{elsarticle}
%% \documentclass[final,5p,times]{elsarticle}
%% \documentclass[final,5p,times,twocolumn]{elsarticle}

%% For including figures, graphicx.sty has been loaded in
%% elsarticle.cls. If you prefer to use the old commands
%% please give \usepackage{epsfig}

%% The amssymb package provides various useful mathematical symbols
\usepackage{amssymb}
%% The amsthm package provides extended theorem environments
%% \usepackage{amsthm}

%% The lineno packages adds line numbers. Start line numbering with
%% \begin{linenumbers}, end it with \end{linenumbers}. Or switch it on
%% for the whole article with \linenumbers.
%% \usepackage{lineno}

\journal{XXXXX}

\begin{document}

\begin{frontmatter}

%% Title, authors and addresses

%% use the tnoteref command within \title for footnotes;
%% use the tnotetext command for theassociated footnote;
%% use the fnref command within \author or \address for footnotes;
%% use the fntext command for theassociated footnote;
%% use the corref command within \author for corresponding author footnotes;
%% use the cortext command for theassociated footnote;
%% use the ead command for the email address,
%% and the form \ead[url] for the home page:
%% \title{Title\tnoteref{label1}}
%% \tnotetext[label1]{}
%% \author{Name\corref{cor1}\fnref{label2}}
%% \ead{email address}
%% \ead[url]{home page}
%% \fntext[label2]{}
%% \cortext[cor1]{}
%% \affiliation{organization={},
%%             addressline={},
%%             city={},
%%             postcode={},
%%             state={},
%%             country={}}
%% \fntext[label3]{}

\title{Predicting the Quality of CDRH3 Antibody Loop Structural Models}

%% use optional labels to link authors explicitly to addresses:
%% \author[label1,label2]{}
%% \affiliation[label1]{organization={},
%%             addressline={},
%%             city={},
%%             postcode={},
%%             state={},
%%             country={}}
%%
%% \affiliation[label2]{organization={},
%%             addressline={},
%%             city={},
%%             postcode={},
%%             state={},
%%             country={}}

\author[inst1]{Lilian M Denzler}

\affiliation[inst1]{organization={Department of Structural and Molecular Biology, 
University College of London},%Department and Organization
            addressline={Gower Street}, 
            city={London},
            postcode={WC1E 6BT}, ,
            country={UK}}

\author[inst1]{Andrew CR Martin}

\begin{abstract}
%% Text of abstract
Therapeutic antibodies have shown an unprecedented pace of development and have brought new hope for the treatment of numerous diseases. The bioinformatic tools for modelling antibody structures have become invaluable for antibody engineering and the development of therapeutic antibodies. The antigen-binding site consists of six hypervariable loops, also known as the Complementary Determining Regions (CDR), all of which can be modelled with adequate accuracy, except for one. It remains markedly difficult to model the third CDR loop of the antibody heavy chain. The CDRH3 differs in length, has far greater sequence variability and has such a great structural diversity that modelling it is considerably harder. 
Many sophisticated approaches for antibody modelling, such as the abYmod software, have been developed. Although such efforts have improved prediction accuracy the results for the CDRH3 loop are still inconsistent and require further improvement. Providing a confidence score for the structure predictions would aid in differentiating well-modelled structures from incorrectly modelled structures, giving the abYmod user a clearer understanding of the generated model reliability. 
We present a model quality predictor, combining domain knowledge with machine learning techniques to predict the accuracy of CDRH3 models generated by antibody modelling software such as abYmod. The newly developed predictor scored a Mathews Correlation Coefficient of 0.99, and can thus be described as highly reliable. The predictor is made available at \url{http://www.bioinf.org.uk/abs/qualiloop/}.

\end{abstract}

%%Graphical abstract
%%\begin{graphicalabstract}
%%\includegraphics{grabs}
%%\end{graphicalabstract}

%%Research highlights
%%\begin{highlights}
%%\item Research highlight 1
%%\item Research highlight 2
%%\end{highlights}

%%\begin{keyword}
%% keywords here, in the form: keyword \sep keyword
%%keyword one \sep keyword two
%% PACS codes here, in the form: \PACS code \sep code
%%\PACS 0000 \sep 1111
%% MSC codes here, in the form: \MSC code \sep code
%% or \MSC[2008] code \sep code (2000 is the default)
%%\MSC 0000 \sep 1111
%%\end{keyword}

\end{frontmatter}

%% \linenumbers

%% main text

@@

\label{sec:sample1}

@@

In order to rationally design therapeutic antibodies, knowledge of their structure is essential. The acquired structural information can be used to increase binding affinity to a target of interest, predicting both the exact binding site and the antibody stability as well as assessing immunogenicity \cite{Abhinandan2007}. As experimental structure determination is very costly and time consuming, computational predictions of an antibody’s structure are used to streamline the process.  

@@

\begin{figure}[h!]
  \includegraphics[scale=1]{loopdist.png}
  \centering
  \caption { Distribution of CDR-H3 loop lengths in mouse (left), human (centre) and cow (right). Mouse and human antibody CDRH3s have a unimodal, yet almost normal distribution with a range of ca. 4-28 and 4-38 respectively. The length cut-off for CDRH3s depicted above is 30 amino acids. Bovine antibodies with significantly longer CDR-H3 loops than depicted here exist, reaching a length of 67 amino acids and above. \cite{Wong2015}}
  \label{fig:loopdist}
\end{figure}

@@

Several different approaches for generating three-dimensional structure models from antibody sequences exist such as RosettaAntibody \cite{Sircar2009},\cite{Sivasubramanian2009}, ABodyBuilder\cite{Leem2016}, PIGSPro\cite{Lepore2017} and abYmod, UCL’s in-house software developed by Prof. Martin. One of the most used methods is RosettaAntibody, which implements template selection and ab initio CDR-H3 loop modelling using loop fragments and employing specific angle restraints which bias the conformational space towards so-called ‘kinked’ loops\cite{Schoeder2021},\cite{Weitzner2017}. In contrast, ABodyBuilder uses a database search algorithm (FREAD\cite{Choi2010}) for CDR loop modelling. 

@@

 Using these mentioned modelling methods, framework regions can generally be predicted with great accuracy (with better than 1Å RMSD\cite{Almagro2014}), as one can often find a very similar structure for the homology modelling process.  However, the CDR loops are not as easily predicted due to their great diversity. If the canonical conformation of CDR loops 1-5 can be identified, they too can be modelled rather well, often within 1Å RMSD, for CDR-H3 loops the average is usually above 3Å \cite{Almagro2011}. 
ABodyBuilder is a modelling server that provides the user with a confidence score for each region (e.g.CDR-H2) of the antibody model. The given score is the probability that a specific region (e.g. CDR-H2) will be will be modeled within a specific RMSD threshold \cite{Leem2016}}. Thus, it can be used to obtain an expected RMSD value for a given probability (default 75\%). For the CDR-H3 this score is calculated as a function of the loop length.  The confidence scorer is described as robust, but less accurate in the case of CDR loops due to the lack of data.\cite{Leem2016} ABLooper also provides a confidence metric for the CDR-H3 loop model, which is estimated by the diversity of a set of predicted conformations for the same loop\cite{Abanades2022}.However, it remains unclear whether a high prediction diversity score points towards loops with multiple conformations or a low quality model. Furthermore, it remains unclear how well the generated diversity score reflects model quality\cite{Abanades2022}.

Modelling the H3 loop is a hurdle for in silico development of therapeutic antibodies. Currently, there is no definite, reliable way to determine how accurate a generated structural model is within the H3 region. Therefore, the aim of this project will be to produce a user-friendly predictor of H3 model quality. The predictor will give the user an RMSD-range in Ångstöms, in which the generated model lies with a high probability. Making such a confidence score available via the web interface of the in-house modelling software abYmod is a future goal. Such a score is not provided by most modelling programs and would thus be a novel addition.  This information can guide the user in the antibody engineering process. The user has the choice to determine whether the model is to be used in the intended way, or whether the model should be re-worked.

\section{Results}
\label{sec:sample2}
To asses the predictive power regarding the CDR-H3 loop, the software was tested on a test-set of antibody structures used in the 2014 and 2011 Antibody Modelling Assessments \cite{Almagro2014},\cite{Almagro2011}. As the results depicted in \ref{fig:AMA} show, abYmod achieves results similar to, or better than other modelling programs. However, the outliers with very high RMSD values increase abYmod’s RMSD average. The predictor in this work would aim to identify such outlier models.
The predictive power of any machine learning model is largely dependent on the quality and size of the dataset it was trained on. As this is a non-linear, complex, multi-class classification problem, a substantial amount of data was required. Thus, an extensive, verified dataset of antibody structures called abYbank/AbDb\cite{Almagro2011}, established by Prof. Andrew Martin, was utilised (1924 non-redundant structures). The root-mean-square deviation (RMSD) value, a measure of distance between backbone C-α atoms of superimposed crystal structures and modelled structures, is calculated (see methods). This metric for model quality was used to classify models. 
\begin{figure}[h!]
  \centering
  \includegraphics[scale=0.3]{AMA}
  \caption {RMSD values of the CDR-H3 loop for structures from the Antibody Modelling Assessment I
(2011) and AMAII (2014). abYmod outperforms other modelling software in some instances, but
also has much lower accuracy in few outlier cases. (left: an abYmod structure for C705 could be
generated, yet the RMSD calculation failed. Right: Ab01 is the rabbit antibody PDB:4MA3,
which was excluded in the CDR-H3 modelling stage in AMAII due to difficulties modelling the
overall structure previously. Ab01 is shown for the methods, where generated models were
adequate for RMSD calculation.)}
  \label{fig:AMA}
\end{figure}
The full pipeline for creating the final machine learning model that will predict model quality by giving its RMSD range starts with a feature-set calculation using the antibody sequence. The feature set includes many different attributes linked to sequence, physical characteristics, interactions, etc. within as well as outside of the loop. There is a plethora of information to train our classifier on, including packing quality, protrusion, hydrophobicity, pseudo-energy (see protrusion, hydrophobicity, pseudo-energy (see methods for details). The sequence logo \ref{fig:logo} visualizes amino acid occurrence within the loop sequence, elements of which can be extracted as features \cite{Thomsen2012},\cite{Shaner1993}.

\begin{figure}[h!]
  \includegraphics[scale=0.9]{logo.png}
  \caption {Sequence Logo of the CDRH3 loop sequence. Data on amino acid occurrence taken from http://abymod.abysis.org/. Visualized using Seq2Logo.}
  \label{fig:logo}
\end{figure}
After creating the feature dataset, it is pre-processed (cleaning, scaling, encoding, see methods for details). Structures with a resolution below 4Å were removed given their low quality. Identical whole antibody structures were removed from the dataset, while instances of different antibodies that matched in loop sequence were not removed. Models of some of these structures with the same loop sequence differ significantly. The few large RMSD ranges may stem from low resolution, e.g. the highest datapoint contains a structure with a resolution of 3.00 Å.  Residue differences near the loop may also explain the conformational difference. Some of these structures are complexed while others are not, which may also affect the loop structure.

The dataset was also screened for any models which abYmod created using a template sequence from LoopDB, a database of CDR-H3 like loops from all proteins. This was not the case for any of our structures. 
The target data (i.e. RMSD values) are then transformed from numerical values to nominal values so that they can be used for classification. In order to define these nominal categories, the total RMSD range must be divided into categories. This is done either by creating uniform classes i.e. 1-2Å, 2-3Å (the optimal size of which must be determined), etc. or by creating balanced classes. When creating balanced classes, the upper and lower thresholds of a category are chosen in such a way that each class contains an equal number of instances. This approach is chosen to counteract the skewness of the RMSD distribution. However, this was found to negatively affect the final model’s predictive power. Therefore, uniform classes were used. 
They are also transformed into a set of binary values according to a list of RMSD thresholds. This is done so that binary models can be trained, which will predict the probability e.g. that the model’s RMSD is above 2Å, 2.2Å, 2.4Å, and so on. The number of binary classifiers incorporated into the first layer have a great effect on the final model, the general 
trend being that the more binary classifiers are used, the better the nominal prediction. 
\subsection{Feature Encoding and Selection}
\label{sec:sample2}

As some features are in the form of amino acid codes, these must be encoded before they can be passed to a machine learning model. The encoding strategy often determines how efficiently the model learns and how much information can be extracted. Different strategies were employed to represent followed by BLOSUM62 and NLF encoding. The physiochemical encoding strategy was implemented for all models, being the most effective. However, PCA-3 BLOSUM62, a dimensionality-reduced BLOSUM62 encoding method achieved comparable results. 
Feature selection was conducted to improve the ML model’s learning capacity. A high-dimensional feature dataset bears the risk of introducing excessive noise, facilitating model overfitting and can be responsible for an overall decrease in model performance and stability. Each additionally inputted feature forces the model to handle a more complex task, which consumes excess computational power and time and leads to overfitting of the model.

@@

\subsection{Hyperparameter Optimization}
\label{sec:sample2}
In the process of hyperparameter optimization, the configuration of model parameters which results in best performance is selected. This is usually a computationally expensive and manual procedure.
In an effort to automate this process, a population is defined for each model type, so hyperparameter optimization can be conducted automatically for each model and seamlessly integrated into the full model creation process. Two different methods for hyperparameter optimization were tested. The first is a hybrid approach of randomized search and grid search, the second uses a genetic algorithm for optimization. The genetic algorithm was found to achieve slightly better results and was selected for all models.

\subsection{Model Performance}
\label{sec:sample2}
The overall best final model is composed of several different binary classifiers, with an extreme gradient boosting (XGBoost) top-layer nominal classifier. Features were selected using random forest feature selection. A final MCC value of 0.99 could be achieved for a model using the abYmod log file as input as well as the loop model file itself. This value slightly dropped to 0.92 if no such log file was given. This is mainly due to the fact that the template sequence abYmod used to generate the model is unknown in the latter case. 
The classifier predicts whether a model has an RMSD of below 2Å, between 2-4Å,or above 4Å. These cut-off values were selected based on the observation that abYmod generally produces a model with RMSD below 4Å. Incorrectly modelled structures (\ref{fig:AMA}) may be identified by screening for structures estimated to have an RMSD above 4Å. If a very high-quality model is needed one should also exclude models above 2Å.

\section{Discussion}
The results suggest that our presented classifier can differentiate between well-modelled and less well-modelled CDR-H3 loop structures. An MCC value of 0.99 was achieved, which underlines this ability for accurate discrimination. Different methods for data pre-processing, feature encoding, feature selection and hyperparameter optimization were tested. 
Feature encoding methods that were very high-dimensional (one-hot-encoding, BLOSUM62, NLF) were found to be unfavorable. Dimensionality reduction methods (Principal Component Analysis (PCA), Independent Component Analysis (ICA), projection-based methods e.g. t-SNE) were used on BLOSUM62 encoded matrices, which lead to significant improvement. However, a physiochemical encoding strategy was most effective. 
The selection of features incorporated in the training set seemed to be most important for effective learning. A multitude of methods were tested. No one-fit-for-all method for the different models could be found. However, for our top-layer classifier in our final model recursive feature elimination worked best. 
A set of commonly used machine learning algorithms were tested, and the best ML models were incorporated into the final ensemble model. A stacked model approach (consisting of 23 binary classifiers and a single top-layer nominal classifier) was shown to outperform single ML models.
An MCC value of 0.99 was achieved for a classifier predicting whether an input-model has an RMSD value below 2Å, 2Å-4Å or above 4Å. 
Given that abYbank/AbDb is soon to be expanded by an additional ca. 2000 structures, classifier performance on models without a known abYmod-template sequence may be improved by a larger dataset. 
It is conceivable that the described predictor may also be incorporated in the antibody modelling process as a low-quality filter in the future, flagging certain structures for re-modelling.

 In a future research project residue patterns in correlation with RMSD may be analyzed. Possibly, one might identify certain sequence patterns that make accurate modelling with abYmod more difficult. Furthermore, separate classifiers according to loop length can be built. Given that loop length is the most important determinant of model quality, this approach may yield some insight into the challenges of modelling shorter vs longer loops. 
One could also conduct an analysis of the predictor’s behaviour when abYmod is forced to use LoopDB-based modelling. This might shed light on whether the ML model presented in this paper is biased towards abYmod’s used source of template structures. It would also give an indication of how well the predictor would work in combination with other modelling software.

\bibliographystyle{elsarticle-num} 
\bibliography{references}

\appendix{}
\section{Experimental Procedures}
\subsection{Computing}
All machine learning and, feature selection and hyperparameter optimization algorithms were implemented in Python. The Scikit-learn library was used for training models, the Yellowbrick\cite{Bengfort2021} library was utilized for visualization. All code is available at https://github.com/LilianDenzler/qualiloop . 
The code was run under CentOS 7 on an 8-core virtual machine on an Intel Xeon 4208 CPU with 16Gig RAM.

@@

\subsubsection{Duplicate Screening} Using AbDb’s redundancy information it was ensured that no antibodies were present in the dataset more than once. The dataset is additionally screened for duplicate instances.

@@

\subsubsection{Physiochemical Feature Encoding} In a paper by L. Nanni and A.Lumini a new encoding technique is 
presented which was developed for machine learning classifiers. Many physiochemical properties are calculated and transformed using a non-linear Fisher transform for dimensionality reduction.  A vector of length 19 is produced for each amino acid34.  
In a paper on designing a neural network for predicting the packing angle of the light and heavy variable chain of an antibodies, A.C.R. Martin and K.R. Abhinandan introduce an encoding method that produces a four-dimensional physiochemical feature vector25. The amino acid properties used are 1) the total number of side-chain atoms, 2) the number of side-chain atoms in the shortest path from Cα to the most distal atom, 3) the Eisenberg consensus hydrophobicity\cite{Eisenberg1982}, 4) the charge.

\subsection{Dataset-splitting}
 The final model was evaluated using a test set, separated from the training set at the start in a 30/70 split (lock-box principle) 35. The performance of all individual sub-models of the first layer are determined using stratified K-folds cross-validation (k=10) as the dataset is imbalanced, being skewed towards lower RMSD values. The method is differentiable from K-folds cross validation as it uses stratified sampling instead of random sampling. This ensures each class is represented, as the percentage of samples for each class are preserved. 
A validation set, usually used for testing during the optimization stage will be omitted in favour of stratified K-folds cross-validation (k=10) \cite{Krstajic2014} \cite{Kohavi1995}.

@@

\subsection{Feature Calculations}
\begin{figure}[h!]
  \centering
  \includegraphics[scale=0.49]{feature_table.png}
  \caption {A summary of how different feature values were calculated.}
  \label{fig:feature_table}
\end{figure}

\begin{figure}[h!]
  \includegraphics[scale=0.6]{angle.png}
  \caption {Diagram visualizing the process underlying the protrusion calculation. First, the base residues (i.e H95 and H102, shown as red spheres) of the CDR-H3 (shown in red) are identified. Then, a line is drawn between the two Cα atoms of these residues (yellow dashed line). The distance of the Cα-atom of each residue in the CDR-H3 loop to this line is calculated. The residue which has the greatest distance to the line (shown as green sphere) is outputted as one-letter amino acid code and used as feature. The distance in Å (depicted as yellow arrow) is used as the ‘protrusion’ feature. }
  \label{fig:angle}
\end{figure}

\begin{figure}[h!]
  \includegraphics[scale=0.45]{method.png}
  \caption {Simplified pipeline for creating the final machine learning model that will predict model quality by giving its RMSD range.}
  \label{fig:method}
\end{figure}

%% The Appendices part is started with the command \appendix;
%% appendix sections are then done as normal sections

%% If you have bibdatabase file and want bibtex to generate the
%% bibitems, please use
%%

\end{document}
\endinput
%%
%% End of file `elsarticle-template-num.tex'.

>>>>

\documentclass[12pt]{article}
\bibliographystyle{unsrt}
\usepackage{graphicx}
\usepackage{url}
%% The amssymb package provides various useful mathematical symbols
\usepackage{amssymb}

\title{Predicting the Quality of CDRH3 Antibody Loop Structural Models}
\author{Lilian M.\ Denzler and Andrew C.R.\ Martin\\
Structural and Molecular Biology, Division of Biosciences,\\
University College of London,\\
Gower Street,\\
London WC1E 6BT, UK
}

\begin{document}
\maketitle

\begin{abstract}
  Therapeutic antibodies have shown an unprecedented pace of
  development and have brought new hope for the treatment of numerous
  diseases. The bioinformatic tools for modelling antibody structures
  have become invaluable for antibody engineering and the development
  of therapeutic antibodies. The antigen-binding site consists of six
  hypervariable loops, also known as the Complementary Determining
  Regions (CDR), all of which can be modelled with adequate accuracy,
  except for one. It remains markedly difficult to model the third CDR
  loop of the antibody heavy chain. The CDRH3 differs in length, has
  far greater sequence variability and has such a great structural
  diversity that modelling it is considerably harder.

  Many sophisticated approaches for antibody modelling, such as the
  abYmod software, have been developed. Although such efforts have
  improved prediction accuracy the results for the CDRH3 loop are
  still inconsistent and require further improvement. Providing a
  confidence score for the structure predictions would aid in
  differentiating well-modelled structures from incorrectly modelled
  structures, giving the abYmod user a clearer understanding of the
  generated model reliability.

  We present a model quality predictor, combining domain knowledge
  with machine learning techniques to predict the accuracy of CDRH3
  models generated by antibody modelling software such as abYmod. The
  newly developed predictor scored a Mathews Correlation Coefficient
  of 0.99, and can thus be described as highly reliable. The predictor
  is made available at \url{http://www.bioinf.org.uk/abs/qualiloop/}
\end{abstract}

@@

@@

In order to rationally design therapeutic antibodies, knowledge of
their structure is essential. The acquired structural information can
be used to increase binding affinity to a target of interest,
predicting both the exact binding site and the antibody stability as
well as assessing immunogenicity \cite{Abhinandan2007}. As
experimental structure determination is very costly and time
consuming, computational predictions of an antibody's structure are
used to streamline the process.

@@

\begin{figure}
  \includegraphics[scale=1]{loopdist.eps}
  \centering
  \caption { Distribution of CDR-H3 loop lengths in mouse (left),
    human (centre) and cow (right). Mouse and human antibody CDRH3s
    have a unimodal, yet almost normal distribution with a range of
    ca. 4-28 and 4-38 respectively. The length cut-off for CDRH3s
    depicted above is 30 amino acids. Bovine antibodies with
    significantly longer CDR-H3 loops than depicted here exist,
    reaching a length of 67 amino acids and
    above. \protect\cite{Wong2015}}  
  \label{fig:loopdist}
\end{figure}

@@

Several different approaches for generating three-dimensional
structure models from antibody sequences exist such as RosettaAntibody
\cite{Sircar2009},\cite{Sivasubramanian2009},
ABodyBuilder\cite{Leem2016}, PIGSPro\cite{Lepore2017} and abYmod,
UCL's in-house software developed by Prof. Martin. One of the most
used methods is RosettaAntibody, which implements template selection
and ab initio CDR-H3 loop modelling using loop fragments and employing
specific angle restraints which bias the conformational space towards
so-called ‘kinked' loops\cite{Schoeder2021},\cite{Weitzner2017}. In
contrast, ABodyBuilder uses a database search algorithm
(FREAD\cite{Choi2010}) for CDR loop modelling.

@@

Using these mentioned modelling methods, framework regions can
generally be predicted with great accuracy (with better than 1\AA\
RMSD\cite{Almagro2014}), as one can often find a very similar
structure for the homology modelling process.  However, the CDR loops
are not as easily predicted due to their great diversity. If the
canonical conformation of CDR loops 1-5 can be identified, they too
can be modelled rather well, often within 1\AA\ RMSD, for CDR-H3 loops
the average is usually above 3\AA\cite{Almagro2011}.

ABodyBuilder is a modelling server that provides the user with a
confidence score for each region (e.g.CDR-H2) of the antibody
model. The given score is the probability that a specific region
(e.g. CDR-H2) will be will be modeled within a specific RMSD threshold
\cite{Leem2016}. Thus, it can be used to obtain an expected RMSD
value for a given probability (default 75\%). For the CDR-H3 this
score is calculated as a function of the loop length.  The confidence
scorer is described as robust, but less accurate in the case of CDR
loops due to the lack of data.\cite{Leem2016} ABLooper also provides a
confidence metric for the CDR-H3 loop model, which is estimated by the
diversity of a set of predicted conformations for the same
loop\cite{Abanades2022}.However, it remains unclear whether a high
prediction diversity score points towards loops with multiple
conformations or a low quality model. Furthermore, it remains unclear
how well the generated diversity score reflects model
quality\cite{Abanades2022}.

Modelling the H3 loop is a hurdle for in silico development of
therapeutic antibodies. Currently, there is no definite, reliable way
to determine how accurate a generated structural model is within the
H3 region. Therefore, the aim of this project will be to produce a
user-friendly predictor of H3 model quality. The predictor will give
the user an RMSD-range in {\AA}ngstr\"{o}ms, in which the generated model lies
with a high probability. Making such a confidence score available via
the web interface of the in-house modelling software abYmod is a
future goal. Such a score is not provided by most modelling programs
and would thus be a novel addition.  This information can guide the
user in the antibody engineering process. The user has the choice to
determine whether the model is to be used in the intended way, or
whether the model should be re-worked.

\section{Results}
To asses the predictive power regarding the CDR-H3 loop, the software
was tested on a test-set of antibody structures used in the 2014 and
2011 Antibody Modelling Assessments
\cite{Almagro2014},\cite{Almagro2011}. As the results depicted in
\ref{fig:AMA} show, abYmod achieves results similar to, or better than
other modelling programs. However, the outliers with very high RMSD
values increase abYmod's RMSD average. The predictor in this work
would aim to identify such outlier models.

The predictive power of any machine learning model is largely
dependent on the quality and size of the dataset it was trained on. As
this is a non-linear, complex, multi-class classification problem, a
substantial amount of data was required. Thus, an extensive, verified
dataset of antibody structures called abYbank/AbDb\cite{Almagro2011},
established by Prof. Andrew Martin, was utilised (1924 non-redundant
structures). The root-mean-square deviation (RMSD) value, a measure of
distance between backbone C-$\alpha$ atoms of superimposed crystal structures
and modelled structures, is calculated (see methods). This metric for
model quality was used to classify models.

\begin{figure}
  \centering
  \includegraphics[scale=0.3]{AMA.eps}
  \caption {RMSD values of the CDR-H3 loop for structures from the
    Antibody Modelling Assessment I (2011) and AMAII (2014). abYmod
    outperforms other modelling software in some instances, but also
    has much lower accuracy in few outlier cases. (left: an abYmod
    structure for C705 could be generated, yet the RMSD calculation
    failed. Right: Ab01 is the rabbit antibody PDB:4MA3, which was
    excluded in the CDR-H3 modelling stage in AMAII due to
    difficulties modelling the overall structure previously. Ab01 is
    shown for the methods, where generated models were adequate for
    RMSD calculation.)}
  \label{fig:AMA}
\end{figure}

The full pipeline for creating the final machine learning model that
will predict model quality by giving its RMSD range starts with a
feature-set calculation using the antibody sequence. The feature set
includes many different attributes linked to sequence, physical
characteristics, interactions, etc. within as well as outside of the
loop. There is a plethora of information to train our classifier on,
including packing quality, protrusion, hydrophobicity, pseudo-energy
(see protrusion, hydrophobicity, pseudo-energy (see methods for
details). The sequence logo \ref{fig:logo} visualizes amino acid
occurrence within the loop sequence, elements of which can be
extracted as features \cite{Thomsen2012},\cite{Shaner1993}.

\begin{figure}
  \includegraphics[scale=0.9]{logo.eps}
  \caption {Sequence Logo of the CDRH3 loop sequence. Data on amino
    acid occurrence taken from http://abymod.abysis.org/. Visualized
    using Seq2Logo.}
  \label{fig:logo}
\end{figure}

After creating the feature dataset, it is pre-processed (cleaning,
scaling, encoding, see methods for details). Structures with a
resolution below 4\AA\ were removed given their low quality. Identical
whole antibody structures were removed from the dataset, while
instances of different antibodies that matched in loop sequence were
not removed. Models of some of these structures with the same loop
sequence differ significantly. The few large RMSD ranges may stem from
low resolution, e.g. the highest datapoint contains a structure with a
resolution of 3.00 \AA.  Residue differences near the loop may also
explain the conformational difference. Some of these structures are
complexed while others are not, which may also affect the loop
structure.

The dataset was also screened for any models which abYmod created
using a template sequence from LoopDB, a database of CDR-H3 like loops
from all proteins. This was not the case for any of our structures.

The target data (i.e. RMSD values) are then transformed from numerical
values to nominal values so that they can be used for
classification. In order to define these nominal categories, the total
RMSD range must be divided into categories. This is done either by
creating uniform classes i.e. 1-2\AA, 2-3\AA\ (the optimal size of which
must be determined), etc. or by creating balanced classes. When
creating balanced classes, the upper and lower thresholds of a
category are chosen in such a way that each class contains an equal
number of instances. This approach is chosen to counteract the
skewness of the RMSD distribution. However, this was found to
negatively affect the final model's predictive power. Therefore,
uniform classes were used.

They are also transformed into a set of binary values according to a
list of RMSD thresholds. This is done so that binary models can be
trained, which will predict the probability e.g. that the model's RMSD
is above 2\AA, 2.2\AA, 2.4\AA, and so on. The number of binary
classifiers incorporated into the first layer have a great effect on
the final model, the general trend being that the more binary
classifiers are used, the better the nominal prediction.

\subsection{Feature Encoding and Selection}

As some features are in the form of amino acid codes, these must be
encoded before they can be passed to a machine learning model. The
encoding strategy often determines how efficiently the model learns
and how much information can be extracted. Different strategies were
employed to represent followed by BLOSUM62 and NLF encoding. The
physiochemical encoding strategy was implemented for all models, being
the most effective. However, PCA-3 BLOSUM62, a dimensionality-reduced
BLOSUM62 encoding method achieved comparable results.

Feature selection was conducted to improve the ML model's learning
capacity. A high-dimensional feature dataset bears the risk of
introducing excessive noise, facilitating model overfitting and can be
responsible for an overall decrease in model performance and
stability. Each additionally inputted feature forces the model to
handle a more complex task, which consumes excess computational power
and time and leads to overfitting of the model.

@@

\subsection{Hyperparameter Optimization}
In the process of hyperparameter optimization, the configuration of
model parameters which results in best performance is selected. This
is usually a computationally expensive and manual procedure.

In an effort to automate this process, a population is defined for
each model type, so hyperparameter optimization can be conducted
automatically for each model and seamlessly integrated into the full
model creation process. Two different methods for hyperparameter
optimization were tested. The first is a hybrid approach of randomized
search and grid search, the second uses a genetic algorithm for
optimization. The genetic algorithm was found to achieve slightly
better results and was selected for all models.

\subsection{Model Performance}
The overall best final model is composed of several different binary
classifiers, with an extreme gradient boosting (XGBoost) top-layer
nominal classifier. Features were selected using random forest feature
selection. A final MCC value of 0.99 could be achieved for a model
using the abYmod log file as input as well as the loop model file
itself. This value slightly dropped to 0.92 if no such log file was
given. This is mainly due to the fact that the template sequence
abYmod used to generate the model is unknown in the latter case.

The classifier predicts whether a model has an RMSD of below 2\AA,
between 2-4\AA, or above 4\AA. These cut-off values were selected
based on the observation that abYmod generally produces a model with
RMSD below 4\AA. Incorrectly modelled structures (\ref{fig:AMA}) may
be identified by screening for structures estimated to have an RMSD
above 4\AA. If a very high-quality model is needed one should also
exclude models above 2\AA.

\section{Methods}

\subsection{Computing}
All machine learning and, feature selection and hyperparameter
optimization algorithms were implemented in Python. The Scikit-learn
library was used for training models, the
Yellowbrick\cite{Bengfort2021} library was utilized for
visualization. All code is available at
https://github.com/LilianDenzler/qualiloop .

The code was run under CentOS 7 on an 8-core virtual machine on an
Intel Xeon 4208 CPU with 16Gig RAM.

@@

\subsubsection{Duplicate Screening}
Using AbDb's redundancy information it was ensured that no antibodies
were present in the dataset more than once. The dataset is
additionally screened for duplicate instances.

@@

\subsubsection{Physiochemical Feature Encoding}
In a paper by L. Nanni and A.Lumini a new encoding technique is
presented which was developed for machine learning classifiers. Many
physiochemical properties are calculated and transformed using a
non-linear Fisher transform for dimensionality reduction.  A vector of
length 19 is produced for each amino acid34.  In a paper on designing
a neural network for predicting the packing angle of the light and
heavy variable chain of an antibodies, A.C.R. Martin and
K.R. Abhinandan introduce an encoding method that produces a
four-dimensional physiochemical feature vector25. The amino acid
properties used are 1) the total number of side-chain atoms, 2) the
number of side-chain atoms in the shortest path from C$\alpha$ to the most
distal atom, 3) the Eisenberg consensus
hydrophobicity\cite{Eisenberg1982}, 4) the charge.

\subsection{Dataset-splitting}
The final model was evaluated using a test set, separated from the
training set at the start in a 30/70 split (lock-box principle)
35. The performance of all individual sub-models of the first layer
are determined using stratified K-folds cross-validation (k=10) as the
dataset is imbalanced, being skewed towards lower RMSD values. The
method is differentiable from K-folds cross validation as it uses
stratified sampling instead of random sampling. This ensures each
class is represented, as the percentage of samples for each class are
preserved.

A validation set, usually used for testing during the optimization
stage will be omitted in favour of stratified K-folds cross-validation
(k=10) \cite{Krstajic2014} \cite{Kohavi1995}.

@@

\subsection{Feature Calculations}
\begin{figure}
  \centering
  \includegraphics[scale=0.49]{feature_table.eps}
  \caption {A summary of how different feature values were calculated.}
  \label{fig:feature_table}
\end{figure}

\begin{figure}
  \includegraphics[scale=0.6]{angle.eps}
  \caption {Diagram visualizing the process underlying the protrusion
    calculation. First, the base residues (i.e H95 and H102, shown as
    red spheres) of the CDR-H3 (shown in red) are identified. Then, a
    line is drawn between the two C$\alpha$ atoms of these residues (yellow
    dashed line). The distance of the C$\alpha$-atom of each residue in the
    CDR-H3 loop to this line is calculated. The residue which has the
    greatest distance to the line (shown as green sphere) is outputted
    as one-letter amino acid code and used as feature. The distance in
    \AA\ (depicted as yellow arrow) is used as the ‘protrusion'
    feature. }
  \label{fig:angle}
\end{figure}

\begin{figure}
  \includegraphics[scale=0.45]{method.eps}
  \caption {Simplified pipeline for creating the final machine
    learning model that will predict model quality by giving its RMSD
    range.}
  \label{fig:method}
\end{figure}
